---
title: "City case studies"
output: 
  html_document: 
    df_print: paged
    fig_width: 5
    fig_height: 3.5
---

<hr>
## The size bias of city case studies 
<br>

**Hypothesis**: There are more studies for cities of a larger size.

**Method**: Group cities by population size (>10m, 5-10m, 1-5m, <1m), sum and average the studies for each group.

**Data**: Database of urban papers, population data for cities from GEA.

**Results**: Hypothesis is correct for average articles per city. *Note that the majority of cities studies in the database are unallocated (2/3) due to missing population data.* These unallocated cities average very low numbers of studies, but do include some very large ones, e.g. New York City, Stockholm, Santiago, Chicago, Washington D.C., Delhi.
<br>


```{r include=FALSE}

rm(list = ls())
library(ggplot2)
library(dplyr)
library(tidyr)
library(magrittr)
library(reshape2)
library(svglite)

rm(ctop)
load(file = "Data/city_studies.RData")
load(file = "Data/city_data_pnas.RData")

ctop <- left_join(ctop,cpnas, by = c("vars" = "cities"))
ctop <- ctop %>% mutate(population=population*1000)
ctop <- ctop %>%
  group_by(population_category = cut(ctop$population,breaks = c(100e6,10e6,5e6,1e6,1e5),include.lowest=TRUE)) %>%
  summarise(no_cities=n(),sum_articles=sum(n),mean_articles=mean(n),city_names=paste(vars, collapse=", ")) %>% 
  ungroup() %>% arrange(desc(population_category))

```
```{r echo=FALSE,warning=FALSE, rows.print=4}

ctop %>%
  ggplot(.) +
  geom_bar(aes(x=reorder(population_category,-mean_articles),y=mean_articles),stat="identity",fill="#2b8cbe") +
  scale_x_discrete(labels = c(">10m","5-10m","1-5m","<1m","NA")) +
  xlab("Cities, grouped by population") +
  ylab("Average articles")

library(knitr)
library(kableExtra)
kable(ctop, "html", caption="Urban studies versus population") %>%
  kable_styling() %>%
  scroll_box(height = "300px")
```


<br>
<hr>
## The wealth bias of city case studies 
<br>

**Hypothesis**: There are more studies for wealthier cities.

**Method**: Group cities by income classification (World Bank thresholds:$12,235, $3,956, $1,006), sum and average the studies for each group.

**Data**: Database of urban papers, GDPpc data for cities from GEA.

**Results**: Hypothesis is false for average articles per city income bracket. However, there are 3 times **more** total studies in high income cities, and with a greater diversity of cities (although this may be related to greater GDP data avaialability for these cities - *note again the high number of unallocated city studies*). Beijing and Shanghai account for the majority of studies in the middle income bracket (211 out of 286)
<br>


```{r include=FALSE}

rm(ctop)
load(file = "Data/city_studies.RData")
load(file = "Data/city_data_pnas.RData")

ctop <- left_join(ctop,cpnas, by = c("vars" = "cities"))
ctop <- ctop %>%
  group_by(gdp_pc = cut(ctop$gdp_per_cap,breaks = c(1005,3955,12235,1e6),include.lowest=TRUE)) %>% 
  summarise(no_cities=n(),sum_articles=sum(n),mean_articles=mean(n),city_names=paste(vars, collapse=", ")) %>% 
  ungroup() %>% arrange(desc(gdp_pc))
```
```{r echo=FALSE,warning=FALSE, rows.print=4}

ctop %>%
  ggplot(.) +
  geom_bar(aes(x=reorder(gdp_pc,-no_cities),y=mean_articles),stat="identity",fill="#2b8cbe") +
  scale_x_discrete(labels = c(">$12,235","$3,956-$12,235","$1,006-$3,955","NA")) +
  xlab("Cities, grouped by GDP per capita") +
  ylab("Average articles")

kable(ctop, "html", caption="Urban studies versus GDP") %>%
  kable_styling() %>%
  scroll_box(height = "300px")
```

<br><br>
